---
title: "simpleRegress"
author: "Tanya Claros"
date: "Summer 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26
  
```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).
  
```{r}
chesty <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chesty)
summary.lm(chesty)
```


```{r}
new <- data.frame(CHEST=c(17,22,26))
predict(chesty,new)
predict(chesty,new,se.fit=TRUE)
```
  
  
The adjusted r-square for this model (CHEST) is 0.79; the model for NECK has an adjusted r-square of 0.75. So far the best model is the one based on CHEST.

```{r}
hd <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(hd)
summary.lm(hd)
```

```{r}
new <- data.frame(HEADLEN=c(17,22,26))
predict(hd,new)
predict(hd,new,se.fit=TRUE)
```
  
The adjusted r-square for this model (HEADLEN) is 0.84 which is better than all the models; the adjusted r-square for this model (CHEST) is 0.79; the model for NECK has an adjusted r-square of 0.75. So the best model is the one based on HEADLEN.
 